---
title: "basic_workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basic_workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(phreng)
library(TDA)
library(ripserr)
``` 


This vignette introduces the basic workflow for computing persistence homology in {phreng}. The user 
will need to call two function. The first function creates a specification that outlines which type 
of data object and method for computing persistence are being used. The second function takes 
specification from the first function and the data you are using to compute persistence.

We will first work with raster data.
```{r}
model1 <- PH_raster(filtration = "cubical", engine = "ripserr")
model2 <- PH_raster(filtration = "cubical", engine = "TDA")
model1
model2
```
We can see that this specification chooses what engine or R package will be used to compute the persistence.
It also chooses which filtration the computation will use. There are other options the user can specify such
as maximum dimension of the homology or maximum scale of the filtration. 

We can now compute the persistence.
```{r}
out1 <- compute_persistence(model1, volcano)
out2 <- compute_persistence(model2, volcano)
print(as.data.frame(out1))
print(as.data.frame(out2))
```




