---
title: "Basic `phreng` Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic `phreng` Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(phreng)
```

This vignette introduces the basic workflow for computing persistence homology in `phreng`.

## Choosing an Engine

`phreng` supports two different computational engines, `TDA` and `ripserr`, but they do not both support every filtration. Before creating a persistence specification, users can see which filtrations are available, their inputs, and which engines support them.

```{r}
available_filtrations()
```

This returns a table showing which engines support each currently available filtration.

If you already know which filtration you would like to compute, you can use `phreng` to see which engines support it.

```{r}
available_engines("cubical")

available_engines("vietoris_rips")
```

If an unsupported filtration is requested, the function returns an error directing the user toward `available_filtrations()`.

You can also check which input types are supported for a filtration.

```{r}
available_inputs("cubical")

available_inputs("vietoris_rips")
```

You can also check which classes are supported for a particular filtration and/or input.

```{r}
available_classes(
  filtration = "vietoris_rips",
  input = "point_cloud"
)

available_classes(
  filtration = "alpha"
)

available_classes(
  input = "raster"
)
```

Finally, you can use `recommend_engine()` to choose an appropriate engine.

```{r}
recommend_engine(
  filtration = "vietoris_rips",
  input = "point_cloud"
)
```

## Example Workflow

The user will need to call two functions. The first function creates a *specification* that encapsulates which type of data object and method for computing persistence are being used. The second function takes the specification from the first function and the data for which you are computing persistence.

Two specifications designed for point cloud data will be created.

```{r}
spec1 <- PH_pointcloud(filtration = "vietoris_rips", engine = "ripserr", max_dimension = 2)
spec2 <- PH_pointcloud(filtration = "vietoris_rips", engine = "TDA", max_dimension = 2)
```

This specification chooses which R package and implementation, called the *engine*, will be used to compute the persistence. It also chooses which filtration the computation will implement. There are other options the user can specify such as maximum scale of the filtration or which `TDA` library to be used.

The persistence computation will now be completed. The data set `noisy_circle_points` will be used as the example point cloud data.

```{r}
out1 <- compute_persistence(spec1, phutil::noisy_circle_points)
out2 <- compute_persistence(spec2, phutil::noisy_circle_points)
out1 
out2 
```

This same workflow can be replicated with the other available `PH_<math object>` specifications such as `PH_raster`.

### Equivalent Output with Minor Differences

There may be some minor differences between implementations using a different *engine*. The output below shows the birth and death time of the 1 dimensional homological features.

```{r}
out1$pairs[[2]]
out2$pairs[[2]]
```

The values of the births and deaths are same until the $10^{-8}$ths place. Some minor differences are to be expected due to implementation differences between the packages.

## Updating Specifications

This package is designed using `S7`, and the specifications defined by `PH_<math object>` are `S7` objects. After a specification has been defined, it can be updated using the \@ operator.

```{r}
spec1@max_dimension <- 1
spec2@max_dimension <- 1
```

After updates are made, the calculations can be rerun. Now only homological features of dimension 0 and 1 will be included in the calculation.

```{r}
compute_persistence(spec1, phutil::noisy_circle_points)
compute_persistence(spec2, phutil::noisy_circle_points)
```
